Order Preserving Encryption Schemes: \\ Preventing the Leakage of Distribution\\ 6.857 Final Paper \\ Hrishikesh Joshi, Max Kolysh, Mari Miyachi, John Wang

### 2013-05-09 paper.tex
Clearing up the equivalence theorem.

  \subsection{Expectational Equivalence}

  \emph{Definition:} Two OPE schemes $x,y$ are equivalent if for each plaintext $p_i$ that scheme $x$ encodes in a sequence of plaintexts $P = \{p_i\}_{i=1}^n$, there exists some bijective mapping $f: \mathrm{Z}_y \to \mathrm{Z}_x$ where $Pr[Encode_x(c_i) = \omega] = Pr[f(Encode_y(c_i)) = \omega]$ for all $\omega \in \mathrm{Z}_x$ where $\mathrm{Z}_x$ is the space of all possible OPE encodings for scheme $x$, $c_i = Enc(SK, p_i)$ is the ciphertext corresponding to $p_i$, and $Encode_x(c_i)$ and $Encode_y(c_i)$ are the functions that obtain the OPE encoding for schemes $x$ and $y$ respectively.

  In other words, for two schemes to be equivalent, there must exist a mapping $f$ from all of the possible OPE encodings from scheme $x$ to scheme $y$, where the probability of obtaining an encoding in $x$ is the same as the probability of obtaining the corresponding encoding in $y$ (based on the mapping). This must be the case for all possible encodings.

  This notion of equivalence encompasses a number of useful features. In particular, notice that if $x$ is an OPE encoding scheme which is IND-CPDA secure, and $y$ is equivalent to $x$, then $y$ is also IND-CPDA secure. Due to this notion, it is sufficient to show the equivalence of two schemes in order to link the IND-CPDA security of the schemes (the proof of this property is given in the theorem below). Another observation one can make is that the mapping in the equivalence definition is bijective. Hence, if $x$ is equivalent to $y$, then $y$ is necessarily equivalent to $x$ because one can simply use $f^{-1}$ as the mapping, which necessarily exists because of $f$ being bijective.

\begin{theorem}
 If two OPE schemes $x,y$ are equivalent, then $x$ is IND-CPDA secure if and only if $y$ is IND-CPDA secure.
\end{theorem}
\begin{proof}
  First, we observe that notion of equivalence is symmetric (since $f$ is a bijective mapping which is invertible), so that without loss of generality, we need only show that if $x$ is IND-CPDA secure, then $y$ is as well. So now we can suppose that $x$ is IND-CPDA secure and $x,y$ are equivalent schemes.

@@

  \subsection{Equivalence of DCS and RTOPE}

@@

  \begin{table}
    \begin{tabular}{l | c | c | c | c | c | c}
      Trials & 10 & 20 & 50 & 100 & 200 & 500 \\
      \hline \hline
      1000 & 0.111 & 0.122 & 0.179 & 0.249 & 0.363 & 0.537 \\
      5000 & 0.093 & 0.062 & 0.075 & 0.104 & 0.162 & 0.241 \\
     10000 & 0.089 & 0.059 & 0.055 & 0.071 & 0.106 & 0.176 \\
     50000 & 0.090 & 0.047 & 0.028 & 0.037 & 0.052 & 0.077 \\
    100000 & 0.090 & 0.047 & 0.021 & 0.026 & 0.033 & 0.059 \\
    200000 & 0.091 & 0.047 & 0.021 & 0.018 & 0.022 & 0.040 \\
    500000 & 0.091 & 0.047 & 0.019 & 0.013 & 0.017 & 0.025 \\
   1000000 & 0.090 & 0.047 & 0.019 & 0.011 & 0.011 & 0.017 \\
    \end{tabular}
    \caption{Tests}
  \end{table}

>>>>

  \subsection{Structural Expectation Equivalence}

  \emph{Definition:} Two OPE schemes $x,y$ are \emph{Structural Expectation Equivalent} if for each plaintext $p_i$ that scheme $x$ encodes in a sequence of plaintexts $P = \{p_i\}_{i=1}^n$, there exists some bijective mapping $f: \mathrm{Z}_y \to \mathrm{Z}_x$ where $Pr[Encode_x(c_i) = \omega] = Pr[f(Encode_y(c_i)) = \omega]$ for all $\omega \in \mathrm{Z}_x$ where $\mathrm{Z}_x$ is the space of all possible OPE encodings for scheme $x$, $c_i = Enc(SK, p_i)$ is the ciphertext corresponding to $p_i$, and $Encode_x(c_i)$ and $Encode_y(c_i)$ are the functions that obtain the OPE encoding for schemes $x$ and $y$ respectively.

  In other words, for two schemes to be structurally expectation equivalent, there must exist a mapping $f$ from all of the possible OPE encodings from scheme $x$ to scheme $y$, where the probability of obtaining an encoding in $x$ is the same as the probability of obtaining the corresponding encoding in $y$ (based on the mapping). This must be the case for all possible encodings.

  This notion of structural expectation equivalence encompasses a number of useful features. In particular, notice that if $x$ is an OPE encoding scheme which is IND-CPDA secure, and $y$ is structurally expectation equivalent to $x$, then $y$ is also IND-CPDA secure. Due to this notion, it is sufficient to show the structural expectation equivalence of two schemes in order to link the IND-CPDA security of the schemes (the proof of this property is given in the theorem below). Another observation one can make is that the mapping in the structural expectation equivalence definition is bijective. Hence, if $x$ is structurally expectation equivalent to $y$, then $y$ is necessarily structurally expectation equivalent to $x$ because one can simply use $f^{-1}$ as the mapping, which necessarily exists because of $f$ being bijective.

\begin{theorem}
 If two OPE schemes $x,y$ are structurally expectation equivalent, then $x$ is IND-CPDA secure if and only if $y$ is IND-CPDA secure.
\end{theorem}
\begin{proof}
  First, we observe that notion of equivalence is symmetric (since $f$ is a bijective mapping which is invertible), so that without loss of generality, we need only show that if $x$ is IND-CPDA secure, then $y$ is as well. So now we can suppose that $x$ is IND-CPDA secure and $x,y$ are equivalent schemes.

@@

  \subsection{Structural Expectation Equivalence of DCS and RTOPE}

@@

  \begin{theorem}
    DCS and RTOPE are structurally expectation equivalent as the number of nodes $N$ and repetitions of value $x$, $R(x)$, both go to $\infty$, for all $x$ such that $R(x) > 1$.
  \end{theorem}
  \begin{proof}
    We shall prove this theorem by examining a sequence of moves when looking at the insertion of a repeated value $x$. If we can show that the probability of traversing the tree in a particular sequence $\{m_i \}_{i=1}^k$ is the same in both DCS and RTOPE when both schemes start with the same tree, then the two schemes can be shown to be structurally expectation equivalent.
    Denote $x$ as the repeated value to insert. In DCS, to move right so that $m_i = 1$, one must have $p_i < x + r$. To move right, one must have $p_i > x + r$. Although there exists a sequence of moves from the root to a leaf of the tree which includes every node in the tree, we need only examine the nodes which have a plaintext value equal to $x$ (before the addition of the random bit). This is because the traversal of DCS and RTOPE is the same for nodes which are not repeated (they share the normal binary tree traversal rule). Thus, we know that the moves in these cases will be equivalent and the probability of making the moves on non-repeated nodes will be equal.
    Thus, we can reduce our problem to considering a sequence of moves \{$m_i\}_{i=1}^k$ on the plaintext values which are equal to $x$. If we can show that the probability of making a particular sequence of moves on the repeated values is the same for DCS and RTOPE, then we will be done.
  \end{proof}

  \subsection{Slow Convergence}

  \begin{table}
    \begin{tabular}{l | c | c | c | c | c | c}
      Trials & 10 & 20 & 50 & 100 & 200 & 500 \\
      \hline \hline
      1000 & 0.111 & 0.122 & 0.179 & 0.249 & 0.363 & 0.537 \\
      5000 & 0.093 & 0.062 & 0.075 & 0.104 & 0.162 & 0.241 \\
     10000 & 0.089 & 0.059 & 0.055 & 0.071 & 0.106 & 0.176 \\
     50000 & 0.090 & 0.047 & 0.028 & 0.037 & 0.052 & 0.077 \\
    100000 & 0.090 & 0.047 & 0.021 & 0.026 & 0.033 & 0.059 \\
    200000 & 0.091 & 0.047 & 0.021 & 0.018 & 0.022 & 0.040 \\
    500000 & 0.091 & 0.047 & 0.019 & 0.013 & 0.017 & 0.025 \\
   1000000 & 0.090 & 0.047 & 0.019 & 0.011 & 0.011 & 0.017 \\
    \end{tabular}
    \caption{Average Differences of Expected and Realized Rank}
  \end{table}
